---
title: "New York State Municipal Employee Salaries"
author: "Nergui Ravzanaadii, Varvara Rousseau, Justin Vargas"
output: github_document
---

# Libraries and Settings for Plots

```{r}
library(tidyverse)

library(janitor)

library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d

scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Cleaning and Organizing Data 

```{r}
 payroll_data = 
    read_csv("data/Citywide_Payroll_Data__Fiscal_Year_.csv") %>% 
    clean_names() %>%
    filter(fiscal_year == "2022",
           work_location_borough != "NA",
           work_location_borough != "OTHER",
           work_location_borough != "WASHINGTON DC",
           leave_status_as_of_june_30 != "ON SEPARATION LEAVE",
           leave_status_as_of_june_30 != "SEASONAL",
           title_description != "NA") %>%
       separate(agency_start_date, into = c("month", "day", "start_year"), convert = TRUE) %>% 
      mutate(leave_status = leave_status_as_of_june_30,
             pay_basis = recode(pay_basis, "per Annum" = "Annually"),
             pay_basis = recode(pay_basis, "per Day" = "Daily"),
             pay_basis = recode(pay_basis, "per Hour" = "Hourly"),
             county_name = work_location_borough,
             county_name = recode(county_name, "MANHATTAN" = "NEW YORK"),
             county_name = recode(county_name, "BROOKLYN" = "KINGS",
             job_title = title_description)) %>% 
    select(-payroll_number, -first_name, -last_name, -mid_init, -month, -day, -leave_status_as_of_june_30, -regular_hours, -ot_hours, -regular_gross_paid, -work_location_borough) %>% 
    mutate(county_name = as.factor(county_name),
           pay_basis = as.factor(pay_basis),
           leave_status = as.factor(leave_status),
           job_title = title_description) %>%
  select(-title_description)

head(payroll_data)
```

# Pie Chart for Percentages of Municipal Employees by Leave Status Across All Counties Plot

```{r}
Total_ls = table(pull(payroll_data,leave_status))

Total_ls

labels = c("Active", "Ceased", "On Leave")

piepercent = round(100 * Total_ls / sum(Total_ls), 1)

par(xpd = TRUE) 

pie(Total_ls, labels = paste(labels, sep = " ", piepercent, "%"),
    main = "Percentages of Municipal Employees by Leave Status Across All Counties", col =       viridis(length(Total_ls)))
legend("topright", c("Active", "Ceased", "On Leave"),cex = 0.9, fill = viridis(length(Total_ls)))
```

# Bar Graph for the Number of Municipal Employees by Leave Status in New York State Plot

```{r}
Ls_bar = 
  payroll_data %>%
  group_by(leave_status) %>%
  summarise(
    count = n()
    )

ggplot(Ls_bar, aes(x = leave_status, y = count, fill = leave_status)) + 
geom_bar(position = "dodge", stat = "identity") + 
labs(
    title = "The Number of Municipal Employees by Leave Status in New York State",
    x = "Leave Status", 
    y = "Number of Municipal Employees", 
    fill = "Leave Status" 
    ) + 
scale_y_continuous(labels = scales::comma)
```

# ANOVA Analysis

Null Hypothesis(Ho) = The mean base salary is constant for all counties

Alternative Hypothesis(H1) = The mean base salary is different for all municipalities

```{r}
ms_model = 
  payroll_data %>%
  mutate(county_name = as.factor(county_name)) %>%
  group_by(county_name) %>%
  summarise(ms_mean = base_salary)

res = lm(ms_mean ~ factor(county_name), data = ms_model)

anova_analysis = 
    anova(res) %>% 
    broom::tidy() %>% 
    knitr::kable(caption = "One Way Anova Test for Mean Salary by County")

anova_analysis
```

# Tukey test for Mean Salary by County

```{r}
res1 = aov(ms_mean ~ factor(county_name), data = ms_model)

tukey_ms = TukeyHSD(res1)

tukey_ms

tukey_table = 
    tukey_ms %>% 
    broom::tidy() %>%
    knitr::kable(caption = "Tukey Test for Mean Salary by County")

tukey_table
```

# Mean Base Salary By New York State Counties Plot

```{r}
mean_base_salary_plot =
  payroll_data %>%
  group_by(county_name) %>%
  summarize(
  mean_base_salary = mean(base_salary, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(county_name, mean_base_salary), y = mean_base_salary, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
      labels = scales::comma) +
  labs(
    x = "County Names",
    y = "Mean Base Salary",
    title = "The Average Base Salary for Municipal Employees of New York State By County",
    fill = "New York State County Names"
  ) + 
  theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust = 0.50)) 

mean_base_salary_plot
```
 
# Total Other Pay by New York Counties Plot

```{r}
median_tibble =
  payroll_data %>%
  group_by(county_name) %>%
  summarise(median_total_other_pay = median(total_other_pay))

median_other_pay_plot =
  payroll_data %>%
  left_join(median_tibble, by = "county_name") %>%
  ggplot(aes(x = reorder(county_name, median_total_other_pay), y = total_other_pay, fill = county_name)) +
  geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(-20000, 40000),
    breaks = seq(-20000, 40000, by = 10000)) +
  labs(
    x = "County Names",
    y = "Median Total Other Types of Pay",
    title = "The Median Total Other Pay for Municipal Employees of New York State By County",
    fill = "New York State County Names"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) 

median_other_pay_plot
```

# Total Number of Job Titles by New York State Counties Plot

```{r}
number_job_titles_plot =
  payroll_data %>%
  group_by(county_name, job_title) %>%
  summarise(count = n()) %>%
  ungroup() %>% 
  group_by(county_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(county_name, count), y = count, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
      labels = scales::comma,
    limits = c(0, 1500),
    breaks = seq(0, 1500, by = 250)) +
  labs(
    x = "County Names",
    y = "Number of Job Titles",
    title = "The Number of Job Titles Held By Municipal Employees of New York State By County",
    fill = "New York State County Names"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

number_job_titles_plot
```

# Median Overtime Paid By New York State Counties Plot

```{r}
overtime_tibble = 
    payroll_data %>% 
    group_by(county_name) %>% 
    summarise(median_total_ot_paid = median(total_ot_paid))

median_ot_paid_plot = 
    payroll_data |> 
    left_join(overtime_tibble, by = "county_name") |> 
    ggplot(aes(x = reorder(county_name, median_total_ot_paid ), y = total_ot_paid,
               fill = county_name)) +
    geom_boxplot() +
     scale_y_continuous(
      labels = scales::comma,
    limits = c(-5000, 32500),
    breaks = seq(-5000, 32500, by = 5000)) +
    labs(
    x = "County Names",
    y = "Total Overtime Paid",
    title = "Total Overtime Paid By Each County",
    fill = "New York State County Names"
  ) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 


median_ot_paid_plot

```

# Median Base Salary By Pay basis By New York State Counties (Annually) Plot

```{r}
pay_basis_tibble = 
    payroll_data %>%
    filter(pay_basis == "Annually") |> 
    group_by(county_name) %>% 
    summarise(median_base_salary = median(base_salary))

Median_base_salary_plot_annually =
    payroll_data %>%
    filter(pay_basis == "Annually") |> 
    left_join(pay_basis_tibble, by = "county_name") %>%
    ggplot(aes(x = reorder(county_name, median_base_salary), y = base_salary, 
               fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(20000, 150000),
    breaks = seq(20000, 150000, by = 10000)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Annual Median Base Salary For Each County", 
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

Median_base_salary_plot_annually
```

# Median Base Salary By Pay basis By New York State Counties (Daily) Plot

```{r}
pay_basis_daily_tibble = 
    payroll_data %>%
    filter(pay_basis == "Daily") |> 
    group_by(county_name) |> 
    summarise(median_base_salary_daily = median(base_salary))

Median_base_salary_plot_daily =
    payroll_data %>%
    filter(pay_basis == "Daily") |> 
    left_join(pay_basis_daily_tibble,  by = "county_name") %>%
    ggplot(aes(x = reorder(county_name, median_base_salary_daily) , 
               y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0,625),
    breaks = seq(0, 625, by = 50)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Daily Median Base Salary For Each County", 
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

Median_base_salary_plot_daily
```

# Median Base Salary By Pay basis By New York State Counties (Hourly) Plot

```{r}
pay_basis_hourly_tibble= 
    payroll_data %>%
    filter(pay_basis == "Hourly") |> 
    group_by(county_name) %>%
    summarise(median_base_salary_hourly = median(base_salary))

Median_base_salary_plot_hourly =
    payroll_data %>%
    filter(pay_basis == "Hourly") |> 
    left_join(pay_basis_hourly_tibble,  by = "county_name") %>%
    ggplot(aes(x = reorder(county_name, median_base_salary_hourly), 
               y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0,95),
    breaks = seq(0,95, by = 10)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Hourly Median Base Salary For Each County",
     fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

Median_base_salary_plot_hourly
```

# Median Base Salary By Pay basis By New York State Counties (Prorated) Plot

```{r}
pay_basis_prorated_tibble = 
    payroll_data %>%
    filter(pay_basis == "Prorated Annual") |> 
    group_by(county_name) %>%
    summarise(median_base_salary_prorated = median(base_salary))  
    
Median_base_salary_plot_prorated =
    payroll_data %>%
    filter(pay_basis == "Prorated Annual") |>
    left_join(pay_basis_prorated_tibble,  by = "county_name") %>%   
    ggplot(aes(x = reorder(county_name, median_base_salary_prorated), 
               y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(5000,85000),
    breaks = seq(5000, 85000, by = 10000)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Prorated Annual Median Base Salary For Each County",
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


Median_base_salary_plot_prorated
```

# Plot for Top 10 Job Titles Throughout the New York State Counties

```{r}
job_frequency =
payroll_data %>%
    group_by(county_name, job_title) %>%
    summarise(job_frequency = n()) %>%
    arrange(desc(job_frequency)) %>%
    ungroup()

job_frequency_data = 
    job_frequency %>%
    top_n(job_frequency, n = 10) %>%
    mutate(
     job_title = recode(job_title, "TEACHER- PER SESSION" = "TEACHER PER SESSION"),
     job_title = recode(job_title, "ANNUAL ED PARA" = "EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "TEACHER SPECIAL EDUCATION" = "SPECIAL EDUCATION TEACHER"),
     job_title = recode(job_title, "TEACHER-GENERAL ED" = "GENERAL EDUCATION TEACHER"),
     job_title = recode(job_title, "SUBSTITUTE ED PARA" = "SUBSTITUTE EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "F/T SCHOOL AIDE" = "SCHOOL AIDE"))
    
job_frequency_plot =
    job_frequency_data %>%
    ggplot(aes(x = reorder(job_title, job_frequency), y = job_frequency, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") + 
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0, 120000),
    breaks = seq(0, 120000, by = 10000)) +
  labs(
    x = "Top 10 Job Titles",
    y = "Frequency of Job Titles",
    title = "The Top 10 Job Titles Held By Municipal Employees Throughout the Counties in New York State",
    fill = "New York State County Names"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

job_frequency_plot
```

# The Top Five Job Titles Held by Municipal Employees in Six New York State Counties Plot

```{r}
top5_freq = 
  payroll_data %>%
   filter(county_name %in% c("NEW YORK", "KINGS", "BRONX", "QUEENS", "RICHMOND","ALBANY")) %>%
   group_by(county_name, job_title) %>%
   summarise(number_of_people = n()) %>%
   arrange(desc(number_of_people)) %>%
   slice(1:5)

top5_freq_n = 
  top5_freq %>%
   mutate(
     job_title = as.factor(job_title),
     job_title = recode(job_title, "TEACHER- PER SESSION" = "TEACHER PER SESSION"),
     job_title = recode(job_title, "ANNUAL ED PARA" = "EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "TEACHER SPECIAL EDUCATION" = "SPECIAL EDUCATION TEACHER"),
     job_title = recode(job_title, "ASST DIR OF INTERGVNMENTAL RELTNS FOR THE ALBANY OFFICE" = "ASSISTANT DIRECTOR OF INTERNATIONAL RELATIONS"),
     job_title = recode(job_title, "EMERGENCY MEDICAL SPECIALIST-EMT" = "EMT"))
 
top_plot = 
  ggplot(top5_freq_n, aes(x = reorder(job_title, number_of_people), y = number_of_people)) + geom_point(aes(color = county_name), alpha = 0.5) + 
scale_y_continuous(
  labels = scales::comma,
  limits = c(0,120000),
  breaks = seq(0, 120000, by = 15000)) + 
labs(title = "The Top Five Job Titles Held by Municipal Employees in Six New York State Counties",x = "Job Title", y = "Number of Municipal Employees")  + 
  guides(color = guide_legend(title = "New York State County Names")) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

top_plot
```

# Plot for Top 10 Job Titles Throughout the New York County

```{r}
job_frequency_newyork =
payroll_data %>%
    filter(county_name == "NEW YORK") |> 
    group_by(county_name, job_title) %>%
    summarise(job_frequency = n()) %>%
    arrange(desc(job_frequency)) %>%
    ungroup()

job_frequency_newyork_data = 
    job_frequency_newyork %>%
    top_n(job_frequency, n = 10) %>%
    mutate(
     job_title = recode(job_title, "TEACHER- PER SESSION" = "TEACHER PER SESSION"),
     job_title = recode(job_title, "ANNUAL ED PARA" = "EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "TEACHER SPECIAL EDUCATION" = "SPECIAL EDUCATION TEACHER"),
     job_title = recode(job_title, "TEACHER-GENERAL ED" = "GENERAL EDUCATION TEACHER"),
     job_title = recode(job_title, "SUBSTITUTE ED PARA" = "SUBSTITUTE EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "F/T SCHOOL AIDE" = "SCHOOL AIDE"))
    
job_frequency_newyork_plot =
    job_frequency_newyork_data %>%
    ggplot(aes(x = reorder(job_title, job_frequency), y = job_frequency, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") + 
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0, 120000),
    breaks = seq(0, 120000, by = 10000)) +
  labs(
    x = "Top 10 Job Titles",
    y = "Frequency of Job Titles",
    title = "The Top 10 Job Titles Held By Municipal Employees in New York County",
    fill = "New York State County Names"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

job_frequency_newyork_plot
```

# Plot for Top 10 Job Titles Throughout Albany County

```{r}
job_frequency_albany =
payroll_data %>%
    filter(county_name == "ALBANY") |> 
    group_by(county_name, job_title) %>%
    summarise(job_frequency = n()) %>%
    arrange(desc(job_frequency)) %>%
    ungroup()

job_frequency_albany_data = 
    job_frequency_albany %>%
    top_n(job_frequency, n = 10) %>%
    mutate(
     job_title = recode(job_title, "TEACHER- PER SESSION" = "TEACHER PER SESSION"),
     job_title = recode(job_title, "ANNUAL ED PARA" = "EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "TEACHER SPECIAL EDUCATION" = "SPECIAL EDUCATION TEACHER"),
     job_title = recode(job_title, "TEACHER-GENERAL ED" = "GENERAL EDUCATION TEACHER"),
     job_title = recode(job_title, "SUBSTITUTE ED PARA" = "SUBSTITUTE EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "F/T SCHOOL AIDE" = "SCHOOL AIDE"))
    
job_frequency_albany_plot =
    job_frequency_albany_data %>%
    ggplot(aes(x = reorder(job_title, job_frequency), y = job_frequency, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") + 
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0, 5),
    breaks = seq(0, 5, by = 1)) +
  labs(
    x = "Top 10 Job Titles",
    y = "Frequency of Job Titles",
    title = "The Top 10 Job Titles Held By Municipal Employeesin Albany County",
    fill = "New York State County Names"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

job_frequency_albany_plot
```

Report

Data: 

The data that was used for this project is from a NYC Open Data webpage (hyperlink this) entitled “Citywide Payroll Data (Fiscal Year)”.  It was collected with the purpose of understanding how New York City’s budget is utilized on salary and overtime pay for all municipal employees.  However, it is important to note that the data contains information pertaining to municipal employees throughout various counties in New York State.  This source of data is updated annually and was last updated on October 28, 2022.

The raw data consisted of 5,109,775 rows and 17 columns, which were fiscal year, payroll number, the name of the agency, last name of the employee, first name of the employee, middle initial of the employee, the date that the employee started working for the agency, the borough of the employee’s primary work location, the civil service title description, the leave status of the employee at the end of the fiscal year on June 30th, the base salary of the employee, pay basis, the number of regular hours that the employee worked in the fiscal year, the amount paid to the employee for base salary during the fiscal year, overtime hours worked by the employee during the fiscal year, the total overtime pay that was paid to the employee in the fiscal year, and total other pay.  The borough of the employee’s primary work location variable are counties throughout New York State instead of just being the five boroughs in New York City.  The pay basis is whether is the employee was paid by hour, per diem, or annually.  The total other pay variable is any financial compensation in addition to overtime pay and gross salary, which would include lump sums, meal allowance, settlement amounts, differentials, uniform allowance, retroactive pay increases, and bonus pay.  This data does not show salary increases that are received during any one fiscal year.  

The data was read into R as a .csv file using the “read_csv” function.  All of the variable names were then changed to lowercase and an underscore was used between spaces by using the “clean_names” function.  We then filtered the data to only include records from the fiscal year of 2022.  We also filtered the data to remove records containing “NA”, “OTHER”, and “WASHINGTON DC” for the “work_location_borough” variable.  In addition, we also filtered the data to remove records consisting of “ON SEPARATION LEAVE” and “SEASONAL” for the “leave_status_as_of_june_30” variable, and removed records containing “NA” for the “title_description” variable.  The “agency_start_date” variable, which is the date that the employee started at the agency, was separated into three variables, “month”, “day”, and “start_year”.  The variables “leave_status_as_of_june_30”, “work_location_borough”, and “title_description” were duplicated and renamed into “leave_status”, “county_name”, and “job_title”, respectively.  The records containing “per Annum”, “per Day”, and “per Hour” for the pay basis variable were recoded into “Annually”, “Daily”, and “Hourly”, respectively.  The records containing “MANHATTAN” and “BROOKLYN” for the county name variable were recoded into “New York” and “Kings”, respectively.  The variables of payroll number, first name of the employee, last name of the employee, middle initial of the employee, month of the employee’s start date, day of the employee’s start date, the duplicated “leave_status_as_of_june_30”, the duplicated “work_location_borough”, the number of regular hours that the employee worked in the fiscal year, the amount paid to the employee for base salary during the fiscal year, and the overtime hours worked by the employee during the fiscal year were then removed from the data.  The variables for county name, pay basis, and leave status were changed from a character data type to a factor.  The variable for “title_description” was then duplicated and renamed into “job_title”.  The duplicated “title_description” variable was then removed from the data.  The data that was cleaned resulted in 594,234 records and 10 columns.

Exploratory Analysis:

Exploratory analysis for our primary outcome of whether there is a significant difference in the mean base salary of employees in New York State by county in 2022 was conducted by creating a bar graph (Figure 1).  Figure 1 shows the average base salary for municipal employees of New York State by county.  This figure illustrates that Albany county has the highest mean base salary for its municipal employees, while Nassau county has the lowest average base salary and New York county has the second lowest mean base salary.  All of the other counties have similar average base salaries.  However, we hypothesized that New York county would have the highest mean salary as it serves as a major financial center within the United States compared to Albany county, which is the capital of New York State.

We also performed exploratory analysis by investigating the number of job titles that are held by municipal employees in New York State by county in 2022 through a bar graph (Figure 2).  Figure 2 conveys the number of job titles that are held by employees within each of the 16 counties in New York State.  This figure demonstrates that Orange county has the lowest number of job titles, while Albany has the third lowest number of job titles.  It is important to note that each of the five boroughs in New York City are among the counties that have the five highest counts of job titles that are held by municipal employees in 2022: New York county has the highest, which was expected; Queens county has the second highest number; Kings county has the third highest number; Bronx has the fourth highest number; Richmond has the fifth highest number.  We did expect that New York county would have the highest number of job titles held by municipal employees and we also expected that the other boroughs of New York City would also have among the highest counts due to the large size of the populations within each of these counties within New York City.  All of the other counties within New York State may have low frequencies of job titles that are held by municipal workers in 2022, because of their smaller population sizes.

Furthermore, an exploratory analysis was conducted to comprehend the top ten job titles that are held by municipal workers throughout the counties in New York State in 2022 through a bar graph (Figure 3).  Based on Figure 3, the top ten job titles by frequency are Correction Officer, School Aide, Police Officer, Substitute Education Paraprofessional, General Education Teacher, Special Education Teacher, Education Paraprofessional, Election Worker, Teacher, and Teacher Per Session, respectively in ascending order.  The top 9 job titles by frequency were in New York county with Teacher Per Session being the job title with the highest frequency.  The job title with the 10th highest frequency was in Queens county and was a Correction Officer.  We did predict that one of the teacher job titles would be the job title with the highest frequency, but we did not expect that it would be the Teacher Per Session job title.  We also expected that all of the top 10 job titles with the highest frequencies would be in New York County.  Overall, this figure supports some of our predictions and illustrates that the job title with the highest frequency in Queens is a Correction Officer.

An exploratory analysis was performed to understand the median total other pay for municipal employees throughout New York State by county in 2022 through a boxplot (Figure 4).  Figure 4 demonstrates that New York county has the lowest median for total other pay, which is relatively close to 0 dollars, while Schoharie county has the highest median for total other pay, which is relatively above 5000 dollars for municipal workers.  We did not expect New York county to have the lowest median for total other pay, but instead be one of the counties with the highest median for total other pay due to the large population and number of job titles within this county (Figure 2).  All of the other counties have relatively similar medians for total other pay in 2022.  It is also important to note that New York county has the greatest number of outliers, which may be due to the large population within this county.  Based on figure 3, Queens, Bronx, and Kings counties also have high numbers of outliers, which may also be due to the large populations within these counties.  There are counties that do not have any outliers, such as Sullivan, Delaware, Greene, and Orange counties.  This figure also illustrates that there are employees in some of the counties that have negative values for other types of pay, which can include lump sums, meal allowance, settlement amounts, differentials, uniform allowance, retroactive pay increases, and bonus pay.  These other forms of payment would explain why there are negative values for this variable.

