---
title: "New York State Municipal Employee Salaries"
author: "Nergui Ravzanaadii, Varvara Rousseau, Justin Vargas"
output: github_document
---

# Libraries and Settings for Plots

```{r}
library(tidyverse)

library(janitor)

library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d

scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Cleaning and Organizing Data 

```{r}
 payroll_data = 
    read_csv("data/Citywide_Payroll_Data__Fiscal_Year_.csv") %>% 
    clean_names() %>%
    filter(fiscal_year == "2022",
           work_location_borough != "NA",
           work_location_borough != "OTHER",
           work_location_borough != "WASHINGTON DC",
           leave_status_as_of_june_30 != "ON SEPARATION LEAVE",
           leave_status_as_of_june_30 != "SEASONAL",
           title_description != "NA") %>%
       separate(agency_start_date, into = c("month", "day", "start_year"), convert = TRUE) %>% 
      mutate(leave_status = leave_status_as_of_june_30,
             pay_basis = recode(pay_basis, "per Annum" = "Annually"),
             pay_basis = recode(pay_basis, "per Day" = "Daily"),
             pay_basis = recode(pay_basis, "per Hour" = "Hourly"),
             county_name = work_location_borough,
             county_name = recode(county_name, "MANHATTAN" = "NEW YORK"),
             county_name = recode(county_name, "BROOKLYN" = "KINGS",
             job_title = title_description)) %>% 
    select(-payroll_number, -first_name, -last_name, -mid_init, -month, -day, -leave_status_as_of_june_30, -regular_hours, -ot_hours, -regular_gross_paid, -work_location_borough) %>% 
    mutate(county_name = as.factor(county_name),
           pay_basis = as.factor(pay_basis),
           leave_status = as.factor(leave_status),
           job_title = title_description) %>%
  select(-title_description)

head(payroll_data)
```

# Pie Chart for Percentages of Municipal Employees by Leave Status Across All Counties Plot

```{r}
Total_ls = table(pull(payroll_data,leave_status))

Total_ls

labels = c("Active", "Ceased", "On Leave")

piepercent = round(100 * Total_ls / sum(Total_ls), 1)

par(xpd = TRUE) 

pie(Total_ls, labels = paste(labels, sep = " ", piepercent, "%"),
    main = "Percentages of Municipal Employees by Leave Status Across All Counties", col =       viridis(length(Total_ls)))
legend("topright", c("Active", "Ceased", "On Leave"),cex = 0.9, fill = viridis(length(Total_ls)))
```

# Bar Graph for the Number of Municipal Employees by Leave Status in New York State Plot

```{r}
Ls_bar = 
  payroll_data %>%
  group_by(leave_status) %>%
  summarise(
    count = n()
    )

ggplot(Ls_bar, aes(x = leave_status, y = count, fill = leave_status)) + 
geom_bar(position = "dodge", stat = "identity") + 
labs(
    title = "The Number of Municipal Employees by Leave Status in New York State",
    x = "Leave Status", 
    y = "Number of Municipal Employees", 
    fill = "Leave Status" 
    ) + 
scale_y_continuous(labels = scales::comma)
```

# ANOVA Analysis

Null Hypothesis(Ho) = The mean base salary is constant for all counties

Alternative Hypothesis(H1) = The mean base salary is different for all municipalities

```{r}
ms_model = 
  payroll_data %>%
  mutate(county_name = as.factor(county_name)) %>%
  group_by(county_name) %>%
  summarise(ms_mean = base_salary)

res = lm(ms_mean ~ factor(county_name), data = ms_model)

anova_analysis = 
    anova(res) %>% 
    broom::tidy() %>% 
    knitr::kable(caption = "One Way Anova Test for Mean Salary by County")

anova_analysis
```

# Tukey test for Mean Salary by County

```{r}
res1 = aov(ms_mean ~ factor(county_name), data = ms_model)

tukey_ms = TukeyHSD(res1)

tukey_ms

tukey_table = 
    tukey_ms %>% 
    broom::tidy() %>%
    knitr::kable(caption = "Tukey Test for Mean Salary by County")

tukey_table
```

# Mean Base Salary By New York State Counties Plot

```{r}
mean_base_salary_plot =
  payroll_data %>%
  group_by(county_name) %>%
  summarize(
  mean_base_salary = mean(base_salary, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(county_name, mean_base_salary), y = mean_base_salary, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
      labels = scales::comma) +
  labs(
    x = "County Names",
    y = "Mean Base Salary",
    title = "The Average Base Salary for Municipal Employees of New York State By County",
    fill = "New York State County Names"
  ) + 
  theme(axis.text.x = element_text(angle = 80, vjust = 0.5, hjust = 0.50)) 

mean_base_salary_plot
```
 
# Total Other Pay by New York Counties Plot

```{r}
median_tibble =
  payroll_data %>%
  group_by(county_name) %>%
  summarise(median_total_other_pay = median(total_other_pay))

median_other_pay_plot =
  payroll_data %>%
  left_join(median_tibble, by = "county_name") %>%
  ggplot(aes(x = reorder(county_name, median_total_other_pay), y = total_other_pay, fill = county_name)) +
  geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(-20000, 40000),
    breaks = seq(-20000, 40000, by = 10000)) +
  labs(
    x = "County Names",
    y = "Median Total Other Types of Pay",
    title = "The Median Total Other Pay for Municipal Employees of New York State By County",
    fill = "New York State County Names"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) 

median_other_pay_plot
```

# Total Number of Job Titles by New York State Counties Plot

```{r}
number_job_titles_plot =
  payroll_data %>%
  group_by(county_name, job_title) %>%
  summarise(count = n()) %>%
  ungroup() %>% 
  group_by(county_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(county_name, count), y = count, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(
      labels = scales::comma,
    limits = c(0, 1500),
    breaks = seq(0, 1500, by = 250)) +
  labs(
    x = "County Names",
    y = "Number of Job Titles",
    title = "The Number of Job Titles Held By Municipal Employees of New York State By County",
    fill = "New York State County Names"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

number_job_titles_plot
```

# Nergui's Visualization 
# First boxplot for overtime pay by county

```{r}
overtime_tibble = payroll_data %>% 
    group_by(county_name) %>% 
    summarise(median_total_ot_paid = median(total_ot_paid))
    
median_ot_paid_plot = 
    payroll_data |> 
    left_join(overtime_tibble, by = "county_name") |> 
    ggplot(aes(x = reorder(county_name, median_total_ot_paid ), y = total_ot_paid, fill = county_name)) +
    geom_boxplot() +
     scale_y_continuous(
      labels = scales::comma,
    limits = c(-5000, 32500),
    breaks = seq(-5000, 32500, by = 5000)) +
    labs(
    x = "County Names",
    y = "Total Overtime Paid",
    title = "Total Overtime Paid By Each County",
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 


median_ot_paid_plot

```

# Second Boxplot for Median base salary by pay basis by county (Annually)

```{r}
pay_basis_tibble = 
    payroll_data %>%
    filter(pay_basis == "Annually") |> 
    group_by(county_name) %>% 
    summarise(median_base_salary = median(base_salary))

Median_base_salary_plot_annually =
    payroll_data %>%
    left_join(pay_basis_tibble, by = "county_name") %>%
    ggplot(aes(x = reorder(county_name, median_base_salary), y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(20000, 150000),
    breaks = seq(20000, 150000, by = 10000)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Annual Median Base Salary For Each County", 
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

Median_base_salary_plot_annually
```

# Daily 

```{r}
pay_basis_daily_tibble = 
    payroll_data %>%
    filter(pay_basis == "Daily") |> 
    group_by(county_name) |> 
    summarise(median_base_salary_daily = median(base_salary))

Median_base_salary_plot_daily =
    payroll_data %>%
    left_join(pay_basis_daily_tibble,  by = "county_name") %>%
    ggplot(aes(x = reorder(county_name, median_base_salary_daily) , y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0,625),
    breaks = seq(0, 625, by = 50)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Daily Median Base Salary For Each County", 
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

Median_base_salary_plot_daily
```

# Hourly

```{r}
pay_basis_hourly_tibble= 
    payroll_data %>%
    filter(pay_basis == "Hourly") |> 
    group_by(county_name) %>%
    summarise(median_base_salary_hourly = median(base_salary))

Median_base_salary_plot_hourly =
    payroll_data %>%
    left_join(pay_basis_hourly_tibble,  by = "county_name") %>%
    ggplot(aes(x = reorder(county_name, median_base_salary_hourly), y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0,95),
    breaks = seq(0,95, by = 10)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Hourly Median Base Salary For Each County",
     fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

Median_base_salary_plot_hourly
```

# Prorated 

```{r}
pay_basis_prorated_tibble = 
    payroll_data %>%
    filter(pay_basis == "Prorated Annual") |> 
    group_by(county_name) %>%
    summarise(median_base_salary_prorated = median(base_salary))  
    
Median_base_salary_plot_prorated =
    payroll_data %>%
    left_join(pay_basis_prorated_tibble,  by = "county_name") %>%   
    ggplot(aes(x = reorder(county_name, median_base_salary_prorated), y = base_salary, fill = county_name)) +
    geom_boxplot() +
    scale_y_continuous(
      labels = scales::comma,
    limits = c(5000,85000),
    breaks = seq(5000, 85000, by = 10000)) +
    labs(
    x = "County Names",
    y = " Base Salary",
    title = "Prorated Annual Median Base Salary For Each County",
    fill = "New York State County Names"
  ) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


Median_base_salary_plot_prorated
```

# Plot for Top 10 Job Titles Throughout the New York State Counties

```{r}
job_frequency =
payroll_data %>%
    group_by(county_name, job_title) %>%
    summarise(job_frequency = n()) %>%
    arrange(desc(job_frequency)) %>%
    ungroup()

job_frequency_data = 
    job_frequency %>%
    top_n(job_frequency, n = 10) %>%
    mutate(
     job_title = recode(job_title, "TEACHER- PER SESSION" = "TEACHER PER SESSION"),
     job_title = recode(job_title, "ANNUAL ED PARA" = "EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "TEACHER SPECIAL EDUCATION" = "SPECIAL EDUCATION TEACHER"),
     job_title = recode(job_title, "TEACHER-GENERAL ED" = "GENERAL EDUCATION TEACHER"),
     job_title = recode(job_title, "SUBSTITUTE ED PARA" = "SUBSTITUTE EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "F/T SCHOOL AIDE" = "SCHOOL AIDE"))
    
job_frequency_plot =
    job_frequency_data %>%
    ggplot(aes(x = reorder(job_title, job_frequency), y = job_frequency, fill = county_name)) +
  geom_bar(position = "dodge", stat = "identity") + 
    scale_y_continuous(
      labels = scales::comma,
    limits = c(0, 120000),
    breaks = seq(0, 120000, by = 10000)) +
  labs(
    x = "Top 10 Job Titles",
    y = "Frequency of Job Titles",
    title = "The Top 10 Job Titles Held By Municipal Employees Throughout the Counties in New York State",
    fill = "New York State County Names"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

job_frequency_plot
```

# The Top Five Job Titles Held by Municipal Employees in Six New York State Counties Plot

```{r}
top5_freq = 
  payroll_data %>%
   filter(county_name %in% c("NEW YORK", "KINGS", "BRONX", "QUEENS", "RICHMOND","ALBANY")) %>%
   group_by(county_name, job_title) %>%
   summarise(number_of_people = n()) %>%
   arrange(desc(number_of_people)) %>%
   slice(1:5)

top5_freq_n = 
  top5_freq %>%
   mutate(
     job_title = as.factor(job_title),
     job_title = recode(job_title, "TEACHER- PER SESSION" = "TEACHER PER SESSION"),
     job_title = recode(job_title, "ANNUAL ED PARA" = "EDUCATION PARAPROFESSIONAL"),
     job_title = recode(job_title, "TEACHER SPECIAL EDUCATION" = "SPECIAL EDUCATION TEACHER"),
     job_title = recode(job_title, "ASST DIR OF INTERGVNMENTAL RELTNS FOR THE ALBANY OFFICE" = "ASSISTANT DIRECTOR OF INTERNATIONAL RELATIONS"),
     job_title = recode(job_title, "EMERGENCY MEDICAL SPECIALIST-EMT" = "EMT"))
 
top_plot = 
  ggplot(top5_freq_n, aes(x = reorder(job_title, number_of_people), y = number_of_people)) + geom_point(aes(color = county_name), alpha = 0.5) + 
scale_y_continuous(
  labels = scales::comma,
  limits = c(0,120000),
  breaks = seq(0, 120000, by = 15000)) + 
labs(title = "The Top Five Job Titles Held by Municipal Employees in Six New York State Counties",x = "Job Title", y = "Number of Municipal Employees")  + 
  guides(color = guide_legend(title = "New York State County Names")) +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 

top_plot
```
